---
title: "p8105_hw2_tm2680"
author: "Tatini Mal-Sarkar"
date: "9/25/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1
```{r import_transit}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

This dataset `transit_data` contains variables pertaining to the subway, station name, station latitude and longitude, routes, entry, vending, entrance type, and ADA compliance. So far, I have cleaned the variable names, chosen the aforementioned variables, and transformed the character variable `entry` to a logical variable. The dataset `transit_data` contains `r nrow(transit_data)` rows and `r ncol(transit_data)` columns.

There are `r nrow(distinct(transit_data, line, station_name))` distinct stations. There are `r nrow(distinct(filter(transit_data, ada == TRUE), line, station_name))` ADA-compliant stations. `r summarize(transit_data, mean(entry[vending == "NO"]))` of station entrances without vending allow entrance.

```{r reformat_transit}
transit_data =
  gather(transit_data, `route1`, `route2`, `route3`, `route4`, `route5`, `route6`, `route7`, `route8`,`route9`, `route10`, `route11`, key = "route_number", value = "route_name") 
```

`r nrow(distinct(filter(transit_data, route_name == "A"), line, station_name))` distinct stations serve the A train. Of these stations, `r nrow(distinct(filter(transit_data, route_name == "A", ada == TRUE), line, station_name))` are ADA-compliant. 