---
title: "p8105_hw2_tm2680"
author: "Tatini Mal-Sarkar"
date: "9/25/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
devtools::install_github("p8105/p8105.datasets")
```

# Problem 1
```{r import_transit}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

This dataset `transit_data` contains variables pertaining to the subway, station name, station latitude and longitude, routes, entry, vending, entrance type, and ADA compliance. So far, I have cleaned the variable names, chosen the aforementioned variables, and transformed the character variable `entry` to a logical variable. The dataset `transit_data` contains `r nrow(transit_data)` rows and `r ncol(transit_data)` columns. Before gathering, this dataset is not tidy.

There are `r nrow(distinct(transit_data, line, station_name))` distinct stations. There are `r nrow(distinct(filter(transit_data, ada == TRUE), line, station_name))` ADA-compliant stations. `r summarize(transit_data, mean(entry[vending == "NO"]))` of station entrances without vending allow entrance.

```{r reformat_transit}
transit_data =
  gather(transit_data, route1:route11, key = "route_number", value = "route_name") 
```

`r nrow(distinct(filter(transit_data, route_name == "A"), line, station_name))` distinct stations serve the A train. Of these stations, `r nrow(distinct(filter(transit_data, route_name == "A", ada == TRUE), line, station_name))` are ADA-compliant. 

# Problem 2
```{r import_trash_mr}
trash_data_mr = read_excel("data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, 0)))
```

```{r import_trash_2017}
trash_data_2017 = read_excel("data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2017 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  drop_na(month) %>% 
  mutate(year = 2017)
```

```{r import_trash_2016}
trash_data_2016 = read_excel("data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2016 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  drop_na(month) %>% 
  mutate(year = 2016)
```

```{r combine_trash}
trash_data_16_17 = full_join(trash_data_2016, trash_data_2017) %>% 
  mutate(month = month.name[month])
```

Datasets `trash_data_mr` and `trash_data_16_17` describe data collected from the Healthy Harbor initiative. `trash_data_mr` contains 216 observations. Key variables include date, weight, and volume. `trash_data_16_17`contains 20 observations, with far fewer variables. These variables include month, year, and total precipitation. The total precipitation in 2017 was `r sum(trash_data_2017$total)`. The total precipitation in 2016 was `r sum(trash_data_2016$total)`. The median number of sports balls in a dumpster in 2016 was `r median(trash_data_mr$sports_balls)`.

# Problem 3
```{r load_brfss}
library(p8105.datasets)
brfss_data = brfss_smart2010 %>% 
  janitor::clean_names()
```

